---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Load required packages, then load the data
```{r, echo=TRUE}
library(dplyr)
library(lattice) 
activity<-read.csv(file = "activity.csv" )
```

##Calculating the total number of steps per day and plotting the histogram and outputting mean and median.
What is mean total number of steps taken per day?
```{r, echo=TRUE}
stepday<-aggregate(activity$steps, list(activity$date), sum)
stepday<-as.data.frame(stepday)
colnames(stepday)<-c("Day", "Total_Steps")
print(stepday)

hist(stepday$Total_Steps, main="Histogram of Daily Steps", col="magenta", xlab="Steps")

mean(stepday$Total_Steps, na.rm=TRUE)
median(stepday$Total_Steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
``` {r, echo=TRUE}
byinterval<-aggregate(activity$steps, list(activity$interval), mean, na.rm=TRUE)
colnames(byinterval)<-c("Interval", "Mean_Steps")
plot(byinterval,type="l", main="Time Series of Average Steps by Interval",      ylab="steps", xlab="Interval")
byinterval[which.max(byinterval$Mean_Steps), "Interval"]
```


## Imputing missing values, using the mean steps at that time interval. Using looping in merge function. 
```{r, echo=TRUE}

merger<-merge(activity,byinterval, by.x="interval", by.y="Interval", all.x = TRUE)
merger[is.na(merger$steps)!=FALSE,"steps"]<-merger[is.na(merger$steps), "Mean_Steps"]
newdata<-select(merger, steps, date, interval)
newdata.total<-aggregate(newdata$steps, by = list(newdata$date), FUN = sum)
colnames(newdata.total)<-c("Day", "Total_Steps")
hist(stepday$Total_Steps, main="Histogram of Daily Steps", col="green", xlab="Steps")
mean(newdata.total$Total_Steps, na.rm = TRUE)
median(newdata.total$Total_Steps, na.rm = TRUE)
```
So there doesn't appear to be much change in the mean or median after interpolating the NA values

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
newdata$Day<-as.Date(newdata$date, "%Y-%m-%d")
newdata$weekday<-weekdays(newdata$Day)
weekends<-filter(newdata, weekday=="Saturday"|weekday=="Sunday")
weekends$weekend<-"Weekend"
weekdays<-filter(newdata, weekday!="Saturday"& weekday!="Sunday")
weekdays$weekend<-"Weekday"
newdata2<-arrange(rbind(weekends,weekdays), Day)
newdata2total<-aggregate(newdata2$steps, list(newdata2$interval, newdata2$weekend), mean)
colnames(newdata2total)<-c("Interval","Weekend", "Mean_Steps")
attach(newdata2total)
xyplot(Mean_Steps ~ Interval | Weekend, type= "l", main="Average Steps during weekends and weekdays")
```

